---
title: "textoteR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{textoteR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(textoteR)
```

# Why use the textoteR package?

This package makes it easier to convert text corpora from one format to another. The available formats right now are:

- **IRaMuTeQ format** (a single local .txt file with starred tags indicating distinct texts and metadata variables)
- **TXM format** (as many local .txt files as there are texts in the corpus, as well as a metadata.csv file)
- **R tibble** (i.e. a table within the R environment that will contain both the metadata variables and the text variable itself)

# Example: from a TXM corpus

The package is provided with a **small example of TXM corpus** (multiple .txt files + one metadata.csv file): 9 famous fables of La Fontaine.

Here I get the path to the corpus according to where the package is installed:

```{r}
path_to_txm_corpus=system.file("extdata/fables",
                        package="textoteR")
print(path_to_txm_corpus)
```

Here are the files in the directory:
```{r}
list.files(path_to_txm_corpus)
```

Here is how you can convert this corpus into an R tibble:   

```{r}
txm_to_rtibble(from_dir=path_to_txm_corpus)
```

The format of the corpus can also be changed from TXM to IRaMuTeQ through:

```{r}
txm_to_iramuteq(from_dir=path_to_txm_corpus,
                filename="fables_iramuteq.txt")
```

See the first 10 lines of the created file "fables_iramuteq.txt" (removed from local files right afterwards):

```{r}
cat(readLines("fables_iramuteq.txt")[1:10], sep="\n")
file.remove("fables_iramuteq.txt")
```

# Example: from an IRaMuTeQ corpus

The package is provided with a **small example of IRaMuTeQ corpus** (single .txt file with starred tags): 5 speeches pronounced by French President Macron during the COVID-19 crisis in 2020.

```{r}
path_to_iramuteq_corpus=system.file("extdata", package="textoteR")
```


```{r}
iramuteq_to_rtibble(from_dir=path_to_iramuteq_corpus,
                    filename="macron_covid.txt")
```

The format of the corpus can also be changed from IRaMuTeQ to TXM through:

```{r}
iramuteq_to_txm(from_dir=path_to_iramuteq_corpus,
                filename="macron_covid.txt",
                to_dir="macron_covid_corpus")
```

See the content of directory "macron_covid_corpus" (removed from local files right afterwards), and the content of file txt1.txt :

```{r}
list.files("macron_covid_corpus")
cat(readLines("macron_covid_corpus/txt1.txt"))
unlink("macron_covid_corpus",recursive=TRUE)
```

# Example: from an R tibble

The package contains an R data tibble LVtweets, with tweets, that contains both metadata variables and text.

```{r}
head(LVtweets)
```

Here is how you can export such data into an IRaMuTeQ or TXM format:


```{r}
rtibble_to_txm(rtibble=LVtweets,
               to_dir="LVtweets_txm")
list.files("LVtweets_txm")

# remove directory:
unlink("LVtweets_txm", recursive=TRUE)
```


```{r}
rtibble_to_iramuteq(rtibble=LVtweets,
                    filename="LVtweets_ira.txt")
# remove file: 
file.remove("LVtweets_ira.txt")
```

Note that tweets contain a certain number of special characters (e.g. emojis) and links that might cause TXM or IRaMuTeQ imports to fail. Such text data should probably be cleaned in R before conversion to TXM or IRaMuTeQ formats.

